\documentclass{article}

\input{setup/packages}
\input{setup/styling.tex}
\input{setup/commands}
\renewcommand{\L}{\mathcal{L}}
\addbibresource{ref.bib}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}


\title{TMA4183 - Steel Optimization}
\author{}
\date{Spring 2020}

\begin{document}

\maketitle

\section{Introduction}
We consider the optimal control problem for the controlled cooling of steel. In order to model such a cooling process mathematically, we use a linear heat equation as our state equation, given by
\begin{subequations}
   \label{eq:heat}
   \begin{align}
      \rho c_p \theta_t - \nabla \cdot (k \nabla \theta) &= 0 \quad &\text{in } \Omega \times (0,T),\label{eq:heat-in-omega} \\
      -k \frac{\partial \theta}{\partial \nu} &= u(t) (\theta - \theta_w) \quad &\text{on } \Gamma_1 \times (0,T), \label{eq:state-system-bd-1} \\
      -k \frac{\partial \theta}{\partial \nu} &= 0 \quad &\text{on } \Gamma_2 \times (0,T), \label{eq:state-system-bd-2} \\
      \theta(x, 0) &= \theta_0 &\text{in } \Omega. &
   \end{align}
\end{subequations}
Here, $u = u(t) \colon [0, T] \to \mathbb{R}$ is the scalar-valued, time dependent heat transfer coefficient which acts as our unknown control, $c_p$ is the heat capacity, $k$ is thermal conductivity and $\rho$ is the density. The Neumann boundary condition in \eqref{eq:state-system-bd-1} is based on Newton's law of cooling; $\theta_w$ is the temperature of the applied coolant at the boundary. Similarly \eqref{eq:state-system-bd-2} express heat isolation on the boundary $\sigma_0 \times (0,T)$. \bigskip 

The goal is to determine an optimal cooling strategy $u=u(t)$ over the time $[0,T]$ such that $\theta(x, T)$ is ``as close as possible'' to some desired temperature $\theta_d$, where $T$ denotes the final time of the process. In order to model this we introduce a quadratic tracking term into the cost functional. We also require the control $u$ to be \emph{admissible}, i.e. it is possible to apply the found control $u$ in some practical manner.

More formally, we wish to minimize a cost functional of the form
\begin{subequations}
\begin{equation}\label{eq:cost-func}  % Abuse of notation?
   \min J(\theta, u) = \frac{1}{2} \int_\Omega (\theta(x, T) - \theta_d)^2 \mathop{dx} + \frac{\gamma}{2} \int_0^{T} u^2 \mathop{dt}
\end{equation}
subject to
\begin{equation}
      \theta \colon \Omega \times [0, T] \to \mathbb{R} \text{ solves \eqref{eq:heat}},
\end{equation}
and
\begin{equation}
   u \in U_{\mathrm{ad}} := \{ u \in L^{\infty}(0,T): u_a \leq u(t) \leq u_b \text{ for a.e. } t \in [0,T]; u_a < u_b \}.
   \label{eq:box_constraints}
\end{equation}
\end{subequations}
We have defined $Q \coloneqq \Omega \times (0, T)$. Here $\gamma \geq 0$ is a penalizing parameter called Tikhonov regularization parameter, and the last term in $J(\theta, u)$ is a regularizing term which penalize high costs of the control. We have introduced box constraints to limit the amount of applied coolant to the steel slab and henceforth the constraints are restricted to the boundary $\Sigma_1 = \Gamma_1 \times (0,T)$. \bigskip

In this report we shall first develop first-order optimality conditions for \eqref{eq:heat}, secondly in Section 3 we prove existence of an optimal control, then we consider the projected gradient method as our optimisation method. We apply the gradient-method to a simple test problem using a manufactured solution, then we try it on a more realistic scenario, both cases are done in a simple rectangle geometry, representing a steel slab. Finally we consider further improvements to the model, introducing an ODE describing the evolution of the steel microstructure as the steel is heated. We also consider further work to be done, as implementing a Newton's method with primal-dual active set strategy, however due to time constraints we didn't actually finish and start on the implementation of the method. 

\section{Main part}
\input{sections/adjoint-equation}
\input{sections/statesystem}
\input{sections/real_world}
\input{sections/numerical_methods}
\input{sections/numerical}
\input{sections/further_work}
\input{sections/conlusionetc}


\printbibliography

\appendix

\input{sections/appendix_newton}
% \bibliographystyle{alphabetic}
% \bibliography{ref}  

\end{document}
